Online Verification of Commutativity

### 2020-06-30 acmPaper/paper.tex
Use BibTeX!
3.1

Let $P_{all}$ be the set of all paths in G.\xxx[dg]{Do we need this notation?  I don't see it anywhere.  Delete stuff that doesn't show up in the rest of the paper}

The notation $\partial^{+}(p)$ is used to indicate the start node of path p, and similarly $\partial^-(p)$ is the end node of path p. \xxx[dg]{Same for this notation.  I see $\partial$ being useful, but not $\partial^-$} $\partial(p)$ maps to the tuple (start node of p, end node of p).

@@

The diagram commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$; that is, the composition of maps along any path connecting u to v is independent of path choice. Here, equality of functions is defined by some oracle whose specifics will vary by target functions and domain.

@@

To examine the efficacy of our proposed algorithm, we compare it to some potential alternatives.  Specifically, we examine a na\"{i}ve factorial algorithm, which we identify to be a two-flip tolerant path search, and Murota's historical batch solution~\cite{commutative}.

\subsection{Na\"{i}ve algorithm for online addition}

We start with a reference baseline solution for online addition, where the original diagram is path independent and a new edge is added. Pairs that do not involve the new edge must continue to be path independent because they already belonged to the old diagram.   Any implementation can thus verify only parallel pairs that have been affected by the addition of the new edge.

That the original diagram commutes can be used to narrow down the set of pairs to verify.  Cases of pairs where either both paths or neither path includes the new edge must both remain equal with straightforward reasoning.

The cycle checking steps can also be improved upon. In the case where the identity is assumed, only elementary cycles passing through the new edge need to be verified to be equal to the identity. When the identity is not assumed, cycles that include one of the nodes of the new edge must be checked for the new exit or entry point that the new edge becomes (if the cycle includes the source of the new edge then the cycle followed by the new edge must be equal to the new edge alone and if the cycle include the sink then the new edge followed by the cycle must be equal to the new edge).
Additionally, all cycles that include the new edge would need full verification like in the batch algorithm- the cycle would have to be equal to itself composed with itself, and would have to be verified for all its entry and exit points.\xxx[dg]{This paragraph needs to be rewritten to make cycles clear without the identity stuff.  I'm not sure how to do it; I'll come back to it later}

Thus we are left only to verify the pairs where exactly one of the paths includes the new edge.

@@

\xxx[dg]{Insert an introduction paragraph here about what we're solving before jumping into the details}

\xxx[dg]{As we've discussed, I think the existence of an identity function doesn't have to be assumed, since every node "exists" as itself.  I guess another way to put it is that we assume the 0-edge path exists as the "identity" function}
The solution presented here assumes that there is an implicit identity function on every node, as would be the case in most applications. However, a modified version of the algorithm for diagrams without this assumption is presented in the appendix. The modification is similar in spirit to what was described in the description of the naive baselines.

Like with the online baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
The key observation that allows us to improve on the online baseline is that, for a given source and sink pair, only a single parallel pair needs to be verified. 
This is an implication of Theorem \ref{reductionRule}, expanded on later.
Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute. The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair.

@@

An efficient implementation of the algorithm can run in $O(|V|(|V|+|E|))$ time at a $O(|V|^3)$ space cost. In such an implementation, path finding from a given source node to all potential sink nodes would be done in a single $O(|V|+|E|)$ breadth first search and the results memoized in $O(|V|^2)$ space.

@@

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}

\bibitem{commutative} 
Kazuo Murota: {HOMOTOPY BASE OF ACYCLIC GRAPHS - A COMBINATORIAL ANALYSIS OF COMMUTATIVE BY MEANS OF PREORDERED MATROID}. Discrete Applied Mathematics, vol. 17, Issues 1--2, 135--155 (1987).

\end{thebibliography}
\end{document}

>>>>

The notation $\partial(p)$ maps to the tuple (start node of p, end node of p).

@@

The diagram commutes iff f(p)=f(q) $\forall (p,q)\in R_{all}$; that is, the composition of maps along any path connecting u to v is independent of path choice.

The \textsc{Online addition problem}, given a commuting diagram and a new edge, returns whether the diagram commutes.
We make an abstraction - equality of functions is defined by some oracle whose specifics will vary by target functions and domain - and reduce the \textsc{Online addition problem} to the \textsc{Verification set problem}, which given a diagram and a new edge, returns the set of parallel pairs of paths, such that if and only if the members in each pair are equal, then the new graph must commute. 
The output to the \textsc{Online addition problem} can then be obtained as whether equality checking for all pairs succeeds.

@@

To examine the efficacy of our proposed solution to the \textsc{Verification set problem}, we compare it to some potential alternatives.
Specifically, we examine a na\"{i}ve factorial algorithm, a slightly less na\"{i}ve factorial algorithm which we identify to be a two-flip tolerant path search, and Murota's historical batch solution~\cite{commutative}.

\subsection{Very Na\"{i}ve baseline}

We start with the set of all parallel pairs in the diagram (with the new edge added in), and pare it down to be finite by handling cycles.
We find all simple cycles in the diagram, using a procedure like Johnson's algorithm. 
%TODO cite the paper.
We verify for each cycle that a single traversal starting on any node is equal to the identity function by adding pairs of (cycle starting at node, identity function) for each node in the function to the output verification set.
\xxx[AK]{Need cleaner way to put this.}
As cycles must now be the identity, for any pair in the set of all parallel paths, any instance of a cycle can be removed to obtain an equivalent pair with shorter, cycle free paths.
If the shorter pair has equal paths then the paths in the original pair must also be equal to each other.
It is therefore safe to remove all pairs of paths with cycles, leaving only parallel pairs where neither path has a cycle.
Since there are only a finite number of paths in the graph with no cycles, (to the order of n! as a path is an ordering on nodes), and the parallel pairs are a subset of the possibilities generated by choosing a first and second member from this finite set, the resultant set of all parallel pairs must be finite.
%TODO Check if O(n!n) = O(n). What about O((n!)^2)?
The output of the algorithm, then, is the union of the set of cycle verification pairs, and the parallel pairs found by doing an all-acyclic-path search from every node in the diagram with the new edge added.

\subsection{Less Na\"{i}ve baseline}
We further refine the output of the very na\"{i}ve baseline.
The commutativity of the original diagram could be used to narrow down the set of pairs to verify.
Pairs where both paths do not involve the new edge would remain equal (this would apply to cycles too).
Also, pairs where both paths involved the new edge would have to be equal.
To see why this is true, each path could be thought of as consisting of the composition of three segments, the first segment from the common source to the source of the new edge, the second, the new edge itself, and the third, from the sink of the new edge to the common sink of the parallel pair.
The first segment of both pairs would have to be equal because they existed in as parallel pairs in the original diagram, and similarly the third segments would also have to be equal.
The second segments, consisting of the same edge, would also have to be equal. Therefore the composition of these three equal components would be equal. Note that the new edge could only appear once because cycles have already been dealt with and none of the paths under consideration hit a node twice.
Thus we are left only to verify the pairs where exactly one of the paths includes the new edge.
Further, only pairs where the path includes the new edge once need be checked (lemma ~\ref{one_occurence_lemma}).

@@

We present a polynomial time solution to the \textsc{verification set} problem.
Like with the baseline, we do not concern ourselves with parallel pairs where neither or both paths pass through the new edge.
The key observation that allows us to improve on the online baseline is that, for a given source and sink pair, only a single parallel pair needs to be verified. 
This is an implication of Theorem \ref{reductionRule}, expanded on later.
Theorem \ref{verifyingSet} shows that should our selected set of pairs along with cycles passing through the new edge be verified commutative, the entire diagram must commute. The approach is to identify a parallel pair with exactly one path through the edge for each (source, sink) pair.

@@

An efficient implementation of the algorithm can run in $O(|V|^2(|V|+|E|))$ time at a $O(|V|^3)$ space cost. In such an implementation, path finding from a given source node to all potential sink nodes could be done in a single $O(|V|+|E|)$ breadth first search and the results memoized in $O(|V|^2)$ space.

@@

\bibliography{refs}
\bibliographystyle{plain}

\end{document}
